Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-06-22 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

\Abstract{
Understand the maintenance of forests biodiversity structure,
functioning and dynamics is urgent to anticipate their fate in the
global changing context. Communities diversity are assumed to be shaped
by constant regime of disturbance yielding a diversity peak at
intermediate intensity, but for tropical forests this intermediate
disturbance hypothesis (IDH) remains debated as well as the completeness
their resilience in all taxonomic and functional facets after
disturbance. To disentangle the ecological processes driving communities
response to disturbance we analysed the taxonomic and functional
trajectories following a logging and thinning disturbance gradient.
Specifically we examined over 30 years the trajectories of communities
taxonomic richness and evennes and functional composition, diversity and
redundancy based on 7 leaf, stem and life history traits assessing
species ecology. Trajectories highlighted communities taxonomic
resilience, maintaining their initial differences in composition, and
functional resilience through fast functional trajectory common to all
communities in the functional space. The IDH consistently predicted
communities functional trajectories according to the disturbance
intensity, but poorly represented their taxonomic trajectories, blurried
by the alteration of the functional redundancy and the recruitment of
infrequent, shade tolerant species slowed by competitive exclusion
processes. Although communities functioning had recovered after 30 years
their taxonomic diversity and composition in infrequent species and
their functional redundancy remained altered. these results acknowledged
the need of decades-long recovery cycles to ensure communities complete
resilience, and questioned communities maintenance after repeated
disturbance.
}

@@

The large areas covered with tropical forests worldwide hold crucial
environmental, economic and social values. They provide wood and
multiple non-timber forest products, shelter a diversified fauna,
regulate the local climate, and the carbon, water and nutrient cycles,
and ensure cultural and human well-being. The growing demand in forests
products together with current climatic changes increases the pressure
on remaining forests \citep{Gibson2011a, Morales-Hidalgo2015}and
threatens the maintenance and dynamics in space and time of communities
structure, composition and functioning
\citep{Anderson-Teixeira2013, Sist2015}.

In tropical forest, ecological communities are constantly re-shaped by
natural disturbance events that change both the abiotic environment,
through the fluxes of light, heat and water \citep{Goulamoussene2017},
and the biotic interactions like competition among species
\citep{Herault2018}. The cornerstone of tropical forests ecology is to
understand the processes and the determinants of ecosystems response to
disturbance \citep{White2001, Chazdon2003a}. For now, this has been
largely studied through structural parameters, rapid and convenient to
measure, as aboveground biomass, tree height or stem density
\citep{Piponiot2016, Rutishauser2016}. These structural parameters were
thereafter sucessfully modeled, giving important insights into the
maintenance of ecosystems processes and services
\citep{Denslow2000, Blanc2009}. However the response of forests
diversity in tree species remains unclear, albeit it determines the
productivity, stability and functioning of ecosystems
\citep[\citet{Liang2016}]{Tilman2014} and would be most probably
impacted by the changes induced by disturbance \citep{Baraloto2012a}.

In the short-term disturbance demonstrated negligible or even positive
impacts on communities diversity, which have been formalized by the
intermediate disturbance hypothesis (IDH) stating a maximized species
diversity at intermediate disturbance intensity
\citep{Molino2001, Kariuki2006a, Berry2008a}. Still, validations of the
IDH remain scarce in the long term and mainly rely on the analysis of
species richness that gives limited or misleading information on forests
recovery and functioning \citep{Martin2015, Chaudhary2016}. More
complete analysis would besides encompass communities explicit
composition that is crucial for conservation issues, and communities
evenness to reveal the ecological rules structuring communities
\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, a
functional approach accoutning for species biological attributes and
role in the ecosystem directly links ecosystems biodiversity,
functioning and environmental constraints
\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. The
functional trait-based approach focused on major traits related to
species ecology and performance was sucessfully adopted
\citep{Diaz2005, Villeger2008a}, for example to reveal in tropical
rainforests the deterministic processes that foster after disturbance
the fast growing species with efficient resources acquisition
\citep{Molino2001, Haddad2008, Ruger2009}. This basic deterministic
process entails a functional shift from a dominance of ``conservative''
slow-growing species dealing with scarce resources to ``acquisitive''
fast-growing species with rapid and efficient use of abundant resources
\citep{TerSteege2001, Reich2014, Herault2011}. This shift is translated
into consistent trajectories of key functional traits related to
resource acquisition (leaf area, density and chlorophyll content, and
stem specific gravity and bark thickness), tree growth and reproduction
life history traits (seed mass and maximum height)
\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Besides
communities functional composition and diversity, a key aspect of
communties response to disturbance is the changes in functional
redundancy that quantifies the amount of shared trait values among
species \citep{Carmona2016}. High functional redundancy, as in the very
diverse tropical forests \citep{Bellwood2006}, mitigates the impacts of
species removal on ecosystem functioning and determines communities
resilience \citep{Trenbath1999, Elmqvist2003, Diaz2005}. High functional
redundancy also increases the functional overlap among species and would
foster neutral stochastic processes \citep{Gravel2006}.

>>>>

\Abstract{
Understanding the ecological rules underlying the maintenance of tropical forest
functioning and dynamics is urgent to anticipate their fate in the
global changing context. The huge diversity if tropical forests is often assumed to be shaped
by constant regime of disturbance yielding a diversity peak at
intermediate intensity, but for tropical forests this intermediate
disturbance hypothesis (IDH) remains debated ????as well as the completeness
their resilience in all taxonomic and functional facets after
disturbance????. To disentangle the ecological processes driving communities
response to disturbance we analysed the taxonomic and functional
diversity trajectories following a logging and thinning disturbance gradient.
Specifically we examined over 30 years the trajectories of
taxonomic richness and evennes and functional composition, diversity and
redundancy based on 7 leaf, stem and life history traits. Trajectories highlighted communities taxonomic
resilience, maintaining their initial differences in composition, and
functional resilience through fast functional trajectory common to all
communities in the functional space. The IDH consistently predicted
communities functional trajectories according to the disturbance
intensity, but poorly represented their taxonomic trajectories, blurried
by the alteration of the functional redundancy and the recruitment of
infrequent, shade tolerant species slowed by competitive exclusion
processes. Although communities functioning had recovered after 30 years
their taxonomic diversity and composition in infrequent species and
their functional redundancy remained altered. these results acknowledged
the need of decades-long recovery cycles to ensure communities complete
resilience, and questioned communities maintenance after repeated
disturbance.
}

@@

The large areas covered with tropical forests worldwide hold crucial
environmental, economic and social values. They provide wood and
multiple non-timber forest products, shelter a diversified fauna,
regulate the local climate, and the carbon, water and nutrient cycles,
and ensure cultural and human well-being. The growing demand in forests
products together with current climatic changes increases the pressure
on remaining forests \citep{Gibson2011a, Morales-Hidalgo2015} and
threatens the maintenance and dynamics in space and time of communities
structure, composition and functioning \citep{Anderson-Teixeira2013, Sist2015}.

In tropical forest, ecological communities are constantly re-shaped by
natural disturbance events that change both the abiotic environment,
through the fluxes of light, heat and water \citep{Goulamoussene2017},
and the biotic interactions like competition among species
\citep{Herault2018}. One of the cornerstone of tropical forest ecology is to
understand the processes and drivers of ecosystems response to
disturbance \citep{White2001, Chazdon2003a}. For now, this has been
largely studied through forest structural parameters, rapid and convenient to
measure, as aboveground biomass, tree height or stem density
\citep{Piponiot2016, Rutishauser2016}. These structural parameters were
thereafter sucessfully modeled, giving important insights into the
recovery of ecosystems processes and services
\citep{Herault2018}. However the response of forests
diversity is still unclear, albeit it determines the
productivity, stability and functioning of ecosystems
\citep[\citet{Liang2016}]{Tilman2014} and would be most probably
impacted by the changes induced by disturbance \citep{Baraloto2012a}.

In the short-term, moderate disturbance may lead to positive
impacts on communities diversity, which have been formalized by the
intermediate disturbance hypothesis (IDH) stating a maximized species
diversity at intermediate disturbance intensity
\citep{Molino2001, Kariuki2006a, Berry2008a}. Still, validations of the
IDH remain scarce in the long term and mainly rely on the analysis of
species richness that gives limited or misleading information on forests
recovery and functioning \citep{Martin2015, Chaudhary2016}. More
complete analysis would besides encompass communities
composition that is crucial for conservation issues, and abundance distribution
to reveal the ecological rules structuring communities
\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, a
functional approach accounting for species biological attributes and
role in the ecosystem directly links ecosystems biodiversity,
functioning and environmental constraints
\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. The
functional trait-based approach focused on major traits related to
species ecology and performance was sucessfully adopted
\citep{Diaz2005, Villeger2008a}, for example to reveal in tropical
rainforests the deterministic processes that foster after disturbance
the fast growing species with efficient resources acquisition
\citep{Molino2001, Haddad2008, Ruger2009}. This basic deterministic
process entails a functional shift from a dominance of ``conservative''
slow-growing species dealing with scarce resources to ``acquisitive''
fast-growing species with rapid and efficient use of abundant resources
\citep{TerSteege2001, Reich2014, Herault2011}. This shift is translated
into consistent trajectories of key functional traits related to
resource acquisition (leaf area, density and chlorophyll content, and
stem specific gravity and bark thickness), tree growth and reproduction
life history traits (seed mass and maximum height)
\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Besides
communities functional composition and diversity, a last key aspect of
communities response to disturbance is the changes in functional
redundancy that quantifies the amount of shared trait values among
species \citep{Carmona2016}. High functional redundancy, as in the very
diverse tropical forests \citep{Bellwood2006}, mitigates the impacts of
species removal on ecosystem functioning and determines communities
resilience \citep{Trenbath1999, Elmqvist2003, Diaz2005}. High functional
redundancy also increases the functional overlap among species and would
foster neutral stochastic processes \citep{Gravel2006}.
